Component response rate variation underlies the stability of complex systems

### 2020-03-21 notebook/SI.Rmd
Edit the SI so that data are now called from inst instead of the notebook, this will payoff later when making the SI into a vignette

```{r, echo = FALSE}
source(file = "../R/Figures_3_4.R");
source(file = "R_old/sim_mat.R");
source(file = "R_old/plot_figs.R");
source(file = "../R/utilities.R")
```

@@

- [Stability across increasing $S$](#IncrS) 
- [Stability of random ecological networks](#ecological) 
    - [Competitor networks](#competition) 
    - [Mutualist networks](#mutualism) 
    - [Predator-prey networks](#pred-prey) 
- [Sensitivity of connectance (C) values](#connectance)
    - [C = 0.3](#connect3)
    - [C = 0.5](#connect5)
    - [C = 0.7](#connect7)
    - [C = 0.9](#connect9)
- [Large networks of $C = 0.05$ across $S$ and $\sigma$](#sigma)  
    - [$\sigma$ = 0.3](#sigma3)
    - [$\sigma$ = 0.4](#sigma4)
    - [$\sigma$ = 0.5](#sigma5)
    - [$\sigma$ = 0.6](#sigma6)
- [Sensitivity of distribution of $\gamma$](#gam_dist)
- [Stability of structured networks](#structured)
- [Feasibility of complex systems](#Feasibility)
- [Stability given targeted manipulation of $\gamma$ (genetic algorithm)](#ga)
- [Consistency with Gibbs et al. (2018)](#Gibbs)

@@

```{r, echo = FALSE}
dat <- read.csv(file = "sim_results/C_1/random_all.csv");
dat <- dat[,-1]; # Extra row-indicating column removed
```

@@

```{r, echo = FALSE}
cdat <- read.csv(file = "sim_results/ecology/competition_C_1.csv");
mdat <- read.csv(file = "sim_results/ecology/mutualism_C_1.csv");
pdat <- read.csv(file = "sim_results/ecology/pred-prey_C_1.csv");
```

@@

```{r, echo = FALSE}
C3dat <- read.csv(file = "sim_results/C_other/rand_c-0pt3.csv");
C5dat <- read.csv(file = "sim_results/C_other/rand_c-0pt5.csv");
C7dat <- read.csv(file = "sim_results/C_other/rand_c-0pt7.csv");
C9dat <- read.csv(file = "sim_results/C_other/rand_c-0pt9.csv");
```

@@

```{r, echo = FALSE}
C3dat   <- C3dat[,-1];
C3d_tab <- C3dat[,1:7];
kable(C3d_tab);
```

@@

```{r, echo = FALSE}
C5dat   <- C5dat[,-1];
C5d_tab <- C5dat[,1:7];
kable(C5d_tab);
```

@@

```{r, echo = FALSE}
C7dat   <- C7dat[,-1];
C7d_tab <- C7dat[,1:7];
kable(C7d_tab);
```

@@

```{r, echo = FALSE}
C9dat   <- C9dat[,-1];
C9d_tab <- C9dat[,1:7];
kable(C9d_tab);
```

@@

```{r, echo = FALSE}
S3dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt3.csv");
S4dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt4.csv");
S5dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt5.csv");
S6dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt6.csv");
```

@@

```{r, echo = FALSE}
S3dat   <- S3dat[,-1];
S3d_tab <- S3dat[,1:7];
kable(S3d_tab);
```

@@

```{r, echo = FALSE}
S4dat   <- S4dat[,-1];
S4d_tab <- S4dat[,1:7];
kable(S4d_tab);
```

@@

```{r, echo = FALSE}
S5dat   <- S5dat[,-1];
S5d_tab <- S5dat[,1:7];
kable(S5d_tab);
```

@@

```{r, echo = FALSE}
S6dat   <- S6dat[,-1];
S6d_tab <- S6dat[,1:7];
kable(S6d_tab);
```

@@

```{r, echo = FALSE, fig.align = "center"}
swn <- read.csv(file = "sim_results/swn/old/swn.csv"); # Change when new results are in
use <- c(1:6, 19:21);
kable(swn[,use]);
```

In the above, the mean $C$, complexity of $\mathbf{A}$, and complexity of $\mathbf{M}$ are also shown. For similar magnitudes of complexity as in random networks of $\sigma\sqrt{SC} \gtrapprox 1.26$, variation in $\gamma$ typically results in more stable than unstable systems.

@@

```{r, echo = FALSE, fig.align = "center"}
sfn <- read.csv(file = "sim_results/sfn/sfn.csv"); 
use <- c(1:6, 19:21);
kable(sfn[,use]);
```

@@

```{r, echo = FALSE, fig.align = "center"}
csc <- read.csv(file = "sim_results/ecology/cascade_C_1.csv"); 
use <- c(1:5, 18:19);
kable(csc[,use]);
```

@@

```{r, echo = FALSE}
dat      <- read.csv(file = "sim_results/C_1/random_all.csv");
dat      <- dat[,-1]; # Extra row-indicating column removed
feas_tab <- cbind(dat[,1], dat[,8:13]);
colnames(feas_tab)[1] <- "S";
kable(feas_tab);
```

@@

```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
dat <- dat[,-1];
plot_stable_4(dat);
```

@@

```{r, echo = FALSE}
get_top_evo_out <- function(evo_out, size){
    highest <- max(evo_out);
    gammas  <- NULL;
    while(size > 0){
        pos <- which(evo_out == highest);
        len <- length(pos);
        nli <- NULL;
        for(i in 1:len){
            start          <- pos[i] + 1;
            end            <- pos[i] + highest;
            gammas[[size]] <- evo_out[start:end];
            size           <- size - 1;
            if(size == 0){
                break;
            }
        }
        highest <- highest - 1;
    }
    return(gammas);
}
plot_evo_out <- function(evo_out){
    gammas <- get_top_evo_out(evo_out, size = 9);
    for(i in 1:9){
        gammas[[i]] <- 2 * gammas[[i]]; # Account for mistake in return
    }
    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xlim = c(0, 2.1));
    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 2.1));
    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
          text = expression(paste("Component response rate (",gamma,")")));
    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
}
evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
evo_res <- evo_res[,-1];
kable(evo_res);
```

@@

```{r, fig.height = 4, fig.width = 6, echo = FALSE}
evo_out <- scan(file = "sim_results/evolved/evo_out.txt");
plot_evo_out(evo_out);
```

@@

```{r, echo = FALSE}
constant_complexity <- read.csv("sim_results/constant_complexity.csv")
kable(constant_complexity);
```

@@

References
================================================================================

>>>>

```{r, echo = FALSE}
source(file = "../R/Figures_3_4.R");
source(file = "../R/utilities.R")
```

@@

- [Stability across increasing $S$](#IncrS) 
- [Stability of random ecological networks](#ecological) 
    - [Competitor networks](#competition) 
    - [Mutualist networks](#mutualism) 
    - [Predator-prey networks](#pred-prey) 
- [Sensitivity of connectance (C) values](#connectance)
    - [C = 0.3](#connect3)
    - [C = 0.5](#connect5)
    - [C = 0.7](#connect7)
    - [C = 0.9](#connect9)
- [Large networks of $C = 0.05$ across $S$ and $\sigma$](#sigma)  
    - [$\sigma$ = 0.3](#sigma3)
    - [$\sigma$ = 0.4](#sigma4)
    - [$\sigma$ = 0.5](#sigma5)
    - [$\sigma$ = 0.6](#sigma6)
- [Sensitivity of distribution of $\gamma$](#gam_dist)
- [Stability of structured networks](#structured)
- [Feasibility of complex systems](#Feasibility)
- [Stability given targeted manipulation of $\gamma$ (genetic algorithm)](#ga)
- [Consistency with Gibbs et al. (2018)](#Gibbs)
- [Reproducing simulation results](#repr)
- [Literature cited](#ref)

@@

```{r, echo = FALSE}
dat <- read.csv(file = "../inst/extdata/random_all.csv");
```

@@

```{r, echo = FALSE}
cdat <- read.csv(file = "../inst/extdata/competition_C_1.csv");
mdat <- read.csv(file = "../inst/extdata/mutualism_C_1.csv");
pdat <- read.csv(file = "../inst/extdata/pred-prey_C_1.csv");
```

@@

```{r, echo = FALSE}
C3dat <- read.csv(file = "../inst/extdata/rand_c-0pt3.csv");
C5dat <- read.csv(file = "../inst/extdata/rand_c-0pt5.csv");
C7dat <- read.csv(file = "../inst/extdata/rand_c-0pt7.csv");
C9dat <- read.csv(file = "../inst/extdata/rand_c-0pt9.csv");
```

@@

```{r, echo = FALSE}
C3d_tab <- C3dat[,1:7];
kable(C3d_tab);
```

@@

```{r, echo = FALSE}
C5d_tab <- C5dat[,1:7];
kable(C5d_tab);
```

@@

```{r, echo = FALSE}
C7d_tab <- C7dat[,1:7];
kable(C7d_tab);
```

@@

```{r, echo = FALSE}
C9d_tab <- C9dat[,1:7];
kable(C9d_tab);
```

@@

```{r, echo = FALSE}
S3dat <- read.csv(file = "../inst/extdata/sim_SD_0pt3.csv");
S4dat <- read.csv(file = "../inst/extdata/sim_SD_0pt4.csv");
S5dat <- read.csv(file = "../inst/extdata/sim_SD_0pt5.csv");
S6dat <- read.csv(file = "../inst/extdata/sim_SD_0pt6.csv");
```

@@

```{r, echo = FALSE}
S3d_tab <- S3dat[,1:7];
kable(S3d_tab);
```

@@

```{r, echo = FALSE}
S4d_tab <- S4dat[,1:7];
kable(S4d_tab);
```

@@

```{r, echo = FALSE}
S5d_tab <- S5dat[,1:7];
kable(S5d_tab);
```

@@

```{r, echo = FALSE}
S6d_tab <- S6dat[,1:7];
kable(S6d_tab);
```

@@

```{r, echo = FALSE, fig.align = "center"}
swn <- read.csv(file = "../inst/extdata/swn.csv"); 
use <- c(1:6, 19:20, 27);
kable(swn[,use]);
```

In the above, the complexity of $\mathbf{A}$ and $\mathbf{M}$, and the mean $C$, are also shown. For similar magnitudes of complexity as in random networks of $\sigma\sqrt{SC} \gtrapprox 1.26$, variation in $\gamma$ typically results in more stable than unstable systems.

@@

```{r, echo = FALSE, fig.align = "center"}
sfn <- read.csv(file = "../inst/extdata/sfn.csv"); 
use <- c(1:6, 19:21);
kable(sfn[,use]);
```

@@

```{r, echo = FALSE, fig.align = "center"}
csc <- read.csv(file = "../inst/extdata/cascade_C_1.csv"); 
use <- c(1:5, 18:19);
kable(csc[,use]);
```

@@

```{r, echo = FALSE}
dat      <- read.csv(file = "../inst/extdata/random_all.csv");
feas_tab <- c(1, 8:13);
kable(dat[,feas_tab]);
```

@@

```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
evo_dat <- read.csv(file = "../inst/extdata/evo_results.csv");
plot_stables(evo_dat, S_s = 40);
```

@@

```{r, echo = FALSE}
get_top_evo_out <- function(evo_out, size){
    highest <- max(evo_out);
    gammas  <- NULL;
    while(size > 0){
        pos <- which(evo_out == highest);
        len <- length(pos);
        nli <- NULL;
        for(i in 1:len){
            start          <- pos[i] + 1;
            end            <- pos[i] + highest;
            gammas[[size]] <- evo_out[start:end];
            size           <- size - 1;
            if(size == 0){
                break;
            }
        }
        highest <- highest - 1;
    }
    return(gammas);
}
plot_evo_out <- function(evo_out){
    gammas <- get_top_evo_out(evo_out, size = 9);
    for(i in 1:9){
        gammas[[i]] <- 2 * gammas[[i]]; # Account for mistake in return
    }
    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         xlim = c(0, 2.1));
    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 2.1));
    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
         yaxt = "n", xlim = c(0, 2.1));
    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
          text = expression(paste("Component response rate (",gamma,")")));
    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
}
kable(evo_dat);
```

@@

```{r, fig.height = 4, fig.width = 6, echo = FALSE}
evo_out <- scan(file = "../inst/extdata/evo_out.txt");
plot_evo_out(evo_out);
```

@@

```{r, echo = FALSE}
constant_complexity <- read.csv("../inst/extdata/constant_complexity.csv")
kable(constant_complexity);
```

@@

\clearpage

Reproducing simulation results {#repr}
================================================================================

All results in the main text and the literature cited can be reproduced using the [RandomMatrixStability](https://github.com/bradduthie/RandomMatrixStability) R package, which can be downloaded as instructed at the beginning of this Supplemental Information document. The most relevant R functions for reproducing simulations include the following:

1. `rand_gen_var`: Simulates random complex systems and cascade food webs
2. `rand_rho_var`: Simulates random complex systems across a fixed correlation of $\rho = cor(A_{ij}, A_{ji})$
3. `rand_gen_swn`: Simulates randomly generated small-world networks
4. `rand_gen_sfn`: Simulates randomly generated scale-free networks
5. `Evo_rand_gen_var`: Use a genetic algorithm to find stable random complex systems

For the functions 1-4 above, R output will be a table of results. Below describes the headers of this table to more clearly explain what is being reported.

\footnotesize

```{r, echo = FALSE}
header_descriptions <- read.csv("../inst/extdata/header_descriptions.csv")
kable(header_descriptions);
```

\normalsize

Note that output from `Evo_rand_gen_var` only includes the first seven rows of the table above, and `rand_gen_var` does not include $C$ (which can be defind as an argument). All results presented here and in the main text are available in the [inst/extdata](https://github.com/bradduthie/RandomMatrixStability/tree/master/inst/extdata) folder of the [RandomMatrixStability](https://github.com/bradduthie/RandomMatrixStability) R package.

Literature cited {#ref}
================================================================================
